---
title: "Introduction to Econometrics"
subtitle: "Assignment One"
author: "M. Randy Reid"
output:
  pdf_document: default
  html_document: default
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Using the following 5 observations on X and Y, estimate the intercept and regression coefficient
(estimate α and β in yi = α + βxi + εi) using R:

```{r observations}
x <- c(1, 2, 3, 4, 5)
y <- c(2, 2, 3, 3, 6)
```

Report your output - annotate and explain. Provide a summary regression output from R: 

You can also embed plots, for example:

```{r}
#fit a linear model for y on x, where y is the dependent variable and x is the independent variable
#we're examining how changes in x impact y
model <- lm(y ~ x)

#summarize the fit
print(model)
summary(model)

#Intepretation of Results:
#The slope is estimated to be .90 which reflects the change in y with a single unit change in X
#Our intercept is estimated to be 0.50. As such, Y is 0.50 when x is 0. 
#R-Squared is 0.75 meaning that three forths of the variation in Y is attributable to (or explained by) X. 
#Our model shows a positive linear relationship between the variables. 
```

## Question 2
Obtain predicted values and the residuals, and print these along with the values of Y and X.
```{r}
#recalculate residuals from previous model
model_residuals <- resid(model)

#variables and residuals in single table or date frame
q2_table <- data.frame(x=x,y=y,model_residuals=model_residuals)

print(q2_table)
```

## Question 3
If another observation (obs. number 6) will have X=3, what is your best prediction for Y? Can you
calculate the residual for this prediction?
```{r}
#we know that our model is y(hat) = 0.5 + 0.9x and can therefore predict a new value of y given x = 3 
#by substitution
y_estimate <- 0.5 + 0.9*3
print(y_estimate)
```
       
## Question 4
We're using downloaded data (excel format) so I need install the necessary packages and establish a working directory
```{r}
#We're using downloaded data so I need to establish a working directory
setwd('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files')
#import data
h_data <- read.csv('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files/house.csv')
#print for visual inspection
#print(h_data)

#calculate the means and SD for each variable and organize in a data frame
h_averages <- sapply(h_data, mean)
h_sigma <- sapply(h_data, sd)

h_mean_sigma <- data.frame(house_variables = names(h_data), averages = h_averages, SD = h_sigma)
print(h_mean_sigma)
```

##Question 5
Provide scatterplots for sqft x price and bedrms x price (price on the vertical axis) on two separate plots, also plot a histogram of house prices, and a density plot. Report all 4 plots on one plot window - in R, you can use the par(mfrow=c(2,2))
```{r}
setwd('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files')
h_data <- read.csv('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files/house.csv')

par(mfrow=c(2,2))
#plot 1
plot(h_data$SQFT,h_data$PRICE, main = "Sqft against price", ylab = "Price", xlab = "Sqft")
#plot 2
plot(h_data$BEDRMS,h_data$PRICE, main = "Bedroom against price", ylab = "Price", xlab = "Bedrooms")
#plot 3
hist(h_data$PRICE, main = "Histogram of Housing Prices")
#plot 4
dh_price <- density(h_data$PRICE)
plot(dh_price, main = "Density Plot of x", xlab = "x", ylab = "Density", col = "blue", lwd = 2)
```

##Question 6
Use OLS to estimate the equation: pricei = α + β ∗ sqf ti + εi. Report the summary statistics for the
regression results. What are the estimated values of α and β?
```{r}
setwd('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files')
h_data <- read.csv('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files/house.csv')

#OLS
h_model <- lm(h_data$PRICE ~ h_data$SQFT)

#summary stats
summary(h_model)

#pulling coefficients
estimate_intercept <- coef(h_model)[1]
estimate_beta <- coef(h_model)[2]

#print estimates
print(estimate_beta)
print(estimate_intercept)
```

##Question 7
Using the results from the above regression, plot price and sqft (i.e. scatter plot with the regression line) on one graph, and plot the residuals on another graph
```{r}
setwd('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files')
#import data
h_data <- read.csv('/Users/mrr/Desktop/Economics/Intro.Econometrics/Data Files/house.csv')

#OLS
h_model <- lm(h_data$PRICE ~ h_data$SQFT)

#syntax for scatter plot
plot(h_data$PRICE ~ h_data$SQFT, main = "Scatter Plot and Regression Line", xlab = "Sqft", ylab = "Price", pch=15, col="grey", )

#regressionline
abline(h_model,col="blue", lwd=1)

#residuals
h_residuals <- resid(h_model)
plot(h_residuals)
```